---
title: 'HYD298 - Homework #3'
author: "Andrew Brown"
date: "January 25, 2016"
output: pdf_document
---
\subsection{ Fit a normal distribution to the Sebou River dataset}
```{r}
library(car)
foo=read.csv("../HW1/sebou.csv")

#Fit a normal distribution
#normalize data with first and second moment'
xbar_norm=mean(foo$maxQ)
ssd_norm=sd(foo$maxQ)
z.norm<-(foo$maxQ-xbar_norm)/ssd_norm
#plot normalized data versus normal reference line
qqnorm(z.norm)
abline(0,1)
```
Clearly, the normal distribution is inappropriate for the Sebou River dataset. Confirm with a statistical test:
```{r}
#Statistical test to confirm the visually observed misfit
shapiro.test(foo$maxQ)
```
Reject null hypothesis of normally distributed data.

Normal distribution mean:
```{r}
xbar_norm
```
Normal distribution variance:
```{r}
ssd_norm^2
```
Quantiles 1% and 99%:
```{r}
qnorm(mean=xbar_norm,sd=ssd_norm,p=c(0.01,0.99))
```
Note that the 1% quantile is inappropriate for a dataset that necessarily is non-negative. Symmetry of the normal distribution results in tails that correspond to flows below zero.
\newpage
\subsection{ Fit a 2-parameter Lognormal by moments and MLE }
```{r}
#Lognormal
lmaxq=log(foo$maxQ) #log-transform data vector
```

Maximum Likelihood Estimators (Mean):
```{r}
mlemuhat=sum(lmaxq)/length(lmaxq)
mlemuhat
```

Maximum Likelihood Estimators (Variance and Std. Dev.):
```{r}
mlevarhat=sum((lmaxq-mlemuhat)^2)/length(lmaxq)
mlevarhat
sqrt(mlevarhat)
```

Method of Moments (Variance):
```{r}
lvarhat=log(1+ssd_norm^2/xbar_norm^2)
lvarhat
```
Method of Moments (Mean):
```{r}
lmuhat=log(xbar_norm)-0.5*lvarhat
lmuhat
```
The estimates bt MLE and MoM are 

Quantiles 1% and 99%:
```{r}
qlnorm(meanlog=lmuhat,sdlog=sqrt(lvarhat),p=c(0.01,0.99))
```
The lognormal distribution gives more physically realistic quantiles. There is not a 1:1 fit between the data and the lognormal but that is expected given the large uncertainty at the limits of the data range and that the 2parameter lognormal an approximation of a natural gradient. A log transformation produces a more appropriate fit. However, with small sample sizes (i.e. n=43) the results for predictive purposes may vary.
```{r}
z.lnorm<-(lmaxq-lmuhat)/sqrt(lvarhat)
#plot normalized data versus normal reference line
qqnorm(z.lnorm)
abline(0,1)
```